---
title: "Rep.ResearchProject1"
author: "PeterJohnson"
date: "July 21, 2017"
output: html_document
---
Reproducible Research
Project 1
=========================================================
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dev=c("png","html")
```

First, set the working directory

```{r}
setwd("C:/Users/Peter/Desktop/Data Science/Coursera/ReproducibleResearch")
```

#Loading and preprocessing the data

```{r}
url <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

download.file(url,destfil="activity.zip")
unzip("ActivityData.zip", overwrite=TRUE)
data <- read.csv("activity.csv")
```
Now check the data to see that it is downloaded properly

```{r}
head(data)
names(data)
str(data)
```
The data would be easier to use without the missing values
```{r}
good_data <- data[complete.cases(data),]
```


#What is the mean total number of steps taken per day?
You may ignore missing values
1. Calulate the total number of steps taken per day
```{r}
#First convert the date to a factor variable
good_data$date <- factor(good_data$date)
#check the levels
levels(good_data$date)
#aggregate the sums to make a table
table_steps <- aggregate(steps~date,good_data,sum)
table_steps #date is first col, steps is 2nd
```


2. Make a histogram of steps taken per day
```{r}

hist(table_steps$steps,main="Histogram of Total Number of Steps per Day",
     xlab="Total Number of Steps in a day")
#dev.copy(png,"HistNumSteps.png")
#dev.off()
```

3.Calculate the mean and median
```{r}
mean1 <- mean(table_steps$steps)
median1 <- median(table_steps$steps)
```

mean: `mean1`
median:`median1`

#What is the average daily activity pattern
1. Make a time series plot of the 5 min interval and the average number of steps taken, averaged across all days

```{r}
interval_steps <- aggregate(formula= steps~interval,data=good_data,FUN=mean)
plot(interval_steps$interval,interval_steps$steps,type="l",
     main="Average number of steps over all days",
     ylab="Steps Taken", xlab="Interval")
#dev.copy(png,"AVGNumSteps.png")
#dev.off()
```

2. Which 5-minute interval on average contains the maximum number of steps?

```{r}
#first find the index of the max
max_index <- which.max(interval_steps$steps)
interval_steps[max_index,]
```

#Imputing missing values
1. Calculate the total number of missing values in the dataset
 
```{r}
 sum(is.na(data))
```
2. Devise a strategy for filling in all of the missing values in the dataset
 
 We will just replace all NA values with the average value, obviously this is a bad idea and not really representative of the data but thats not the point
 
3. Create a new dataset that is equal to the original, but with the missing data filled in

```{r}
imputed_data <- data
for (i in nrow(imputed_data)){
      if(is.na(imputed_data$steps[i])){
            imputed_data$steps[i] <-mean(table_steps$steps)
      }
}
```
4. Make a histogram of the steps taken each day and calculate the mean and median for this new data set

```{r}

imputed_table_steps <- aggregate(steps~date,imputed_data,sum)
imputed_table_steps #date is first col, steps is 2nd
hist(imputed_table_steps$steps,main="Histogram of Total Number of Steps per Day (Imputed)",
     xlab="Total Number of Steps in a day")
#dev.copy(png,"HistImputedSteps.png")
#dev.off()
imputed_mean <- mean(imputed_table_steps$steps)
imputed_median <- median(imputed_table_steps$steps)
```
mean:`imputed_mean`
median: `imputed_median`

It appears as though imputing the missing data actaully had a minimal effect on the mean and median.

#Are there differences in activity patterns between weekdays and weekends

1. Create a new factor variable with two levels -"weekday" and "weekend"
```{r}
#Extract the day of the week
imputed_data$day <- weekdays(as.Date(imputed_data$date))
imputed_data$day[imputed_data$day %in% c("Saturday","Sunday")] <-"weekend"
imputed_data$day[imputed_data$day != "weekend"] <- "weekday"

#now make into a factor
imputed_data$day <- as.factor(imputed_data$day)
levels(imputed_data$day)
```

2. Make a panel plot containing a time series plot of the 5 min interval and the average number of steps taken averaged across all weekdays or weekend days

```{r}
library(ggplot2)

qplot(interval,steps,data=imputed_data,
      geom=c("line"),xlab="Interval",ylab="Number of steps",ylim=c(0,300)
      )+facet_wrap(~day,ncol=1)
#dev.copy(png,"StepsTimeSeries.png")
#dev.off()
```


